---
title: "Course Project 1"
output: html_document:
  keep_md: true
---

## Code for reading in the dataset and/or processing the data
Reads the csv file and gives a basic summary about the dataset.
```{r}
activity <- read.csv("activity.csv")
head(activity)
str(activity)
summary(activity)

step <- subset(activity, !is.na(activity$steps), na.rm = T)
```

## Histogram of the total number of steps taken each day
Calculates the sum of daily steps and Plots a Histogram for the same.
```{r}
stepsum <- aggregate(steps ~ date, step, sum)
head(stepsum)
hist(stepsum$steps, xlab = "Steps Taken", main = "Total Daily steps", col = "blue")
```

## Mean and median number of steps taken each day
Calculates the mean and median of number of steps taken each day.
```{r}
mn <- round(mean(stepsum$steps), 2)
mn
md <- median(stepsum$steps)
md
```

## Time series plot of the average number of steps taken
Plots a line plot for time over average number of steps taken.
```{r}
averageSteps <- aggregate(steps ~ interval, activity, mean)
plot(averageSteps, type = "l", main = "Average number of steps Taken")
```

## The 5-minute interval that, on average, contains the maximum number of steps
Calculates the maximum number of steps taken on average, in 5 minute interval.
```{r}
maxSteps <- averageSteps[which.max(averageSteps$steps), ]$interval
maxSteps
```

## Code to describe and show a strategy for imputing missing data
Imputing missing data
```{r}
sum(is.na(activity$steps))

impute <- activity
index <- is.na(impute$steps)
avg <- tapply(step$steps, step$interval, mean)
impute$steps[index] <- avg[as.character(impute$interval[index])]
```

## Histogram of the total number of steps taken each day after missing values are imputed
Histogram plot for total number of steps taken each day after imputing missing values.
```{r}
new_stepsum <- aggregate(steps ~ date, impute, sum)
head(new_stepsum)
hist(new_stepsum$steps, col = "blue", xlab = "steps", main = "Total Daily Steps Taken with imputed values")
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
Plot for comparing average number of steps taken per 5 minute interval across weekdays and weekends using lattice plotting system.
```{r}
impute$weekdays <- weekdays(as.Date(impute$date))
impute$DayType <- ifelse(impute$weekdays == "Saturday"|impute$weekdays == "Sunday", "Weekend", "Weekday")
head(impute)

library(lattice)
dayType_stepsum <- aggregate(steps ~ interval+DayType, impute, mean)

xyplot(steps ~ interval | DayType, dayType_stepsum, type = "l", layout = c(1, 2), xlab = "Time Interval", ylab = "Steps", main = "Average No. of steps taken across weekdays and weekends")
```